The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Updated figure 1

\begin{equation*}
\mstar/\mhalo = 2N\left[(\mhalo/M_{1})^{-\beta} + (\mhalo/M_1)^{\gamma}\right]^{-1}.
\end{equation*}
For halo mass $\mhalo$ we use $\mpeak,$ the largest value of $\mvir$ ever attained along the main progenitor branch of the subhalo. The values of the best-fit parameters in \citet{moster_etal13} were fit to a stellar mass function (SMF) with values $\mstar^{\rm MPA-JHU}$ based on the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, which differs from the SMF in our galaxy sample \citep[see, e.g.,][]{bernardi_etal14}. We account for this difference by manually tabulating the median value $\langle\mstar^{\rm Meert+15}\vert\mstar^{\rm MPA-JHU}\rangle$ in logarithmic bins spanning $9 < \log_{10}\mstar^{\rm MPA-JHU}/\msun < 12,$ and applying the median correction to the Monte Carlo realization of the mock galaxy sample. This results in a typical boost of $\sim0.25$ dex at $\mstar^{\rm MPA-JHU}\approx10^{9.75}\msun,$ and $\sim0.4$ dex at $\mstar^{\rm MPA-JHU}\approx10^{11.5}\msun.$

@@

In \S~\ref{subsec:tests1} we show comparisons between the fiducial galaxy profile model described in \S~\ref{sec:model} and our SDSS sample. We demonstrate the sensitivity of galaxy clustering measurements to the post-infall evolution of satellite profiles in \S~\ref{subsec:tests2}, establishing the success of our fiducial model as non-trivial. In \S~\ref{subsec:tests3} we show that even highly idealized measurements of galaxy clustering and lensing are insensitive to assumptions about the relation between halo spin $\halospin$ and size $\rhalf$ of disk-dominated galaxies, implying that large-scale structure observations yield little-to-no constraining power on such assumptions.

@@

\subsubsection{Scaling of $\rhalf$ with $\mstar$}
\label{subsubsec:scaling_relation}

In Figure \ref{fig:scatter_plot} we show the scaling of composite galaxy size $\rhalf$ with $\mstar,$ with separate panels dedicated to subsamples of galaxies divided according to value of $\bt.$ We refer to galaxies with $\bt<0.25$ as ``disk-dominated", $\bt>=0.75$ as ``bulge-dominated", and $0.25<=\bt<0.75$ as ``mixed morphology". We remind the reader that $\rhalf$ and $\mstar$ refer to the scale radius and stellar mass of the {\em entire} galaxy, not of an individual component.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/size_vs_stellar_mass_multipanel_bt_decomposition.pdf}
\caption{
{\bf One-point data used to fit the fiducial model.}
Scattered points show the $\rhalf-\mstar$ relation for SDSS galaxies as measured in \citet{meert_etal15}. Bulge-dominated galaxies are defined in terms of the bulge-to-total stellar mass ratio $\bt>=0.75,$ disk-dominated galaxies $\bt<0.25,$ mixed morphology as $0.25 < \bt < 0.75.$ The black curve in each panel shows the $\rhalf-\mstar$ relation implied by our fiducial model, in which $\rhalfbulge=\abulge\rvir^{\alphabulge}$ and $\rhalfdisk=\adisk\rvir^{\alphadisk},$ with $\adisk=0.014=7\abulge, \alphadisk=1, \alphabulge=5/4,$ and uncorrelated log-normal scatter of $0.2$ dex about these relations.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

>>>>

\begin{equation*}
\langle\mstar/\mhalo\rangle = 2N\left[(\mhalo/M_{1})^{-\beta} + (\mhalo/M_1)^{\gamma}\right]^{-1}.
\end{equation*}
For halo mass $\mhalo$ we use $\mpeak,$ the largest value of $\mvir$ ever attained along the main progenitor branch of the subhalo. The values of the best-fit parameters in \citet{moster_etal13} were fit to a stellar mass function (SMF) with values $\mstar^{\rm MPA-JHU}$ based on the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, which differs from the SMF in our galaxy sample \citep[see, e.g.,][]{bernardi_etal14}. We account for this difference by manually tabulating the median value $\langle\mstar^{\rm Meert+15}\vert\mstar^{\rm MPA-JHU}\rangle$ in logarithmic bins spanning $9 < \log_{10}\mstar^{\rm MPA-JHU}/\msun < 12,$ and applying the median correction to the Monte Carlo realization of the mock galaxy sample. This results in a typical boost of $\sim0.25$ dex at $\mstar^{\rm MPA-JHU}\approx10^{9.75}\msun,$ and $\sim0.4$ dex at $\mstar^{\rm MPA-JHU}\approx10^{11.5}\msun.$

@@

In \S\ref{subsec:tests1} we show comparisons between the galaxy size model described in \S\ref{sec:model} and our SDSS sample. We identify the key ingredients that determine the characteristic $\rhalf-$dependence of galaxy clustering in \S\ref{subsec:tests2}. In so doing, we demonstrate the sensitivity of galaxy clustering measurements to the underlying model assumptions, establishing the success of our model as non-trivial.

@@

In Figure \ref{fig:scatter_plot} we show the scaling of galaxy size $\rhalf$ with $\mstar.$ 
%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/single_component_model_vs_sdss_one_point.pdf}
\caption{
{\bf One-point data used to fit the fiducial model.}
Scattered points show the $\rhalf-\mstar$ relation for SDSS galaxies as measured in \citet{meert_etal15}. The black curve in each panel shows the $\rhalf-\mstar$ relation implied by the model. This figure confirms the findings in \citet{kravtsov13} that a linear relationship between $\rvir$ and $\rhalf,$ convolved against the nonlinear relationships between $\rvir, \mhalo$ and $\mstar,$ correctly predicts the characteristic curvature in the relation $\langle\rhalf\vert\mstar\rangle$ over a wide range in stellar mass.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------
